---
title: "36-315"
author: "Group Number: 19, Names: Julien Naegeli, Alex Frazier, Johnny Wu, Maria Rodriguez: jnaegeli, afrazie1, johnnywu, mar1"
date: "Due Monday"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show

---


```{r}
library(dplyr)
library(readr)
library(ggplot2)

indicators <- read_csv("../Indicators.csv")
indicators <- select(indicators, CountryCode, IndicatorCode, Value, Year)
Countries <- read_csv("../Country.csv")
Countries <- select(Countries, CountryCode, ShortName)
Countries <- mutate(Countries, top_20 = ifelse(ShortName %in% c("Qatar", "Luxembourg", 
                                     "Singapore", "Brunei", "Kuwait", "Norway", "United Arab Emirates", "Hong Kong", "United States", "Switzerland", "Saudi Arabia", "Bahrain", "Netherlands", "Ireland", "Australia", "Austria", "Sweden", "Germany", "Taiwan", "Canada"), "Top 20 Richest Country", "Not Top 20"))
Countries
```

```{r}
#GDP (current LCU)
Variables <- merge(Countries, filter(indicators, IndicatorCode=="NY.GDP.MKTP.CD" &  Year==2012 & Value > 1e+3), by = 'CountryCode')
names(Variables)[names(Variables)=="Value"] <- "GDP"

#Survival rate to the last grade of primary education, both sexes (%)
Variables <- merge(Variables, filter(indicators, IndicatorCode=="SE.PRM.PRSL.ZS" &  Year==2012), by = 'CountryCode')
names(Variables)[names(Variables)=="Value"] <- "Mortality"

GDPMortality <- Variables

#Central government debt, total (% of GDP)
Variables2 <- merge(Countries, filter(indicators, IndicatorCode=="GC.DOD.TOTL.GD.ZS" &  Year==2012), by = 'CountryCode')
names(Variables2)[names(Variables2)=="Value"] <- "Debt"

#Expense (% of GDP)
Variables2 <- merge(Variables2, filter(indicators, IndicatorCode=="GC.XPN.TOTL.GD.ZS" &  Year==2012), by = 'CountryCode')
names(Variables2)[names(Variables2)=="Value"] <- "Expense"


#countsCountry <- indicators %>%
#  group_by(CountryCode, CountryName) %>%
#  summarise(NumIndicators = n_distinct(IndicatorName),
#            NumYears      = n_distinct(Year),
##            FirstYear     = min(Year),
#            LastYear      = max(Year))

#countsIndicate <- indicators %>%
#  group_by(IndicatorCode, IndicatorName) %>%
#  summarise(NumCountries = n_distinct(CountryName),
#            NumYears     = n_distinct(Year),
#            FirstYear    = min(Year),
#            LastYear     = max(Year))

# Note - $ being replaced with dollar to avoid MathJax issue
#counts$IndicatorName <- gsub("\\$", "dollar", counts$IndicatorName)
#print.table(counts)
```

Graphings
```{r}
gdp_mortality_debt_allCountry <- indicators %>% 
  group_by(CountryName, Year, IndicatorName) %>%
  filter(IndicatorName == "GDP (current LCU)" |
         IndicatorName == "Mortality rate, under-5 (per 1,000)" | 
         IndicatorName == "Central government debt, total (% of GDP)" | 
         IndicatorName == "Military expenditure (% of GDP)")
```
```{r}

south_america <- gdp_mortality_debt_allCountry %>%
  filter(CountryName == "Brazil" |
         CountryName == "Argentina" | 
         CountryName == "Paraguay" | 
         CountryName == "Uruguay" |
         CountryName == "Chile" | 
         CountryName == "Bolivia" | 
         CountryName == "Peru" | 
         CountryName == "Ecuador" | 
         CountryName == "Colombia" | 
         CountryName == "Venezuela" | 
         CountryName == "Guyana" | 
         CountryName == "Suriname" )


gdpL <- filter(gdp_mortality_debt_allCountry, IndicatorName == "GDP (current LCU)")
nationalDebt <- filter(gdp_mortality_debt_allCountry, IndicatorName == "Central government debt (% of GDP)")

#gdpL_nationalDebt <- merge(gdpL, nationalDebt, by.x = "CountryName")
jnaegeli_315_theme <-  theme_bw() + 
  theme(
    axis.text = element_text(size = 10, family = "Palatino", 
                                 color = "royalblue4"),
        text = element_text(size = 14, family = "Palatino",
                            face = "bold", color = "grey30"))

#actually GDP^1/4 vs primary education survival rate
ggplot(GDPMortality, aes(x = GDP ^ (1/4), y = Mortality, label = ShortName)) + geom_point(aes(color = top_20)) + 
  labs(
    title = "GDP vs Primary Education Survival Rate",
    x = "GDP (current US Dollar)",
    y = "Survival rate past primary education (%)",
    color = "Country GDP per capita Ranking"
  ) + geom_smooth() + jnaegeli_315_theme + theme(legend.position="bottom")

#+
  #geom_text(data=subset(GDPMortality, GDP ^ (1/4) > 2000), check_overlap = TRUE)


ggplot(Variables2, aes(x = Expense, y = Debt, label = ShortName)) + geom_point(aes(color = top_20)) + 
  labs(
    title = "Government Expenses vs Government Debt",
    x = "Expenses (% of GDP)",
    y = "Central government debt, total (% of GDP)", 
    color = "Country GDP per capita Ranking"
  ) + geom_smooth() +
  geom_text(data=subset(Variables2, Debt > 150)) + jnaegeli_315_theme + theme(legend.position="bottom")
#REMEMBER TO LABEL OUTLIER POINTS AND COLOR THE POINTS AS WELL

#ggplot(nationalDebt, aes(x = Year, y = Value)) + geom_point() + 
#  labs(
#    title = "Government Debt over the years",
#    x = "Central government debt total (% of GDP)",
#    y = "Year"
#  )

#ggplot(gdpL, aes(x = Year, y = Value)) + geom_point() + 
#  labs(
#    title = "GDP over the years",
#    x = "GDP (current LCU)",
#    y = "Year"
#  )
```


1st Graph:
The graph shows survival rate past primary education versus a country's GDP. It seems that there is a general upward trend, as a Country's GDP rises, a larger percentage of the population is able to survive past primary education. This is due in part to better healthcare in richer nations. It is also worth noting that all of the available datapoints for the top 20 richest countries by GDP per capita are near 100% survival rate. 

2nd Graph:
This graph shows a country's expenses versus its debt. As a country spends more, it owes more, increasing its debt. However, the more interesting details lie in the differences, and even lack of differences between top 20 richest nations by GDP per capita and the rest of the countries. All of the outliers and more extreme values come from nations that are not in the top 20. Most of the top 20 nations, while scattered from 0-100 debt-wise, tend to have lower expenses. However, it seems that without the extreme values, the top 20 nations are evenly spaced between the other nations without any visible group or correlation. This suggests that debt and expenses may play a smaller role in a country's wealth than previously thought. 


