---
title: "36-315"
author: "Group Number: 19, Names: Julien Naegeli, Alex Frazier, Johnny Wu, Maria Rodriguez: jnaegeli, afrazie1, johnnywu, mar1"
date: "Due Monday"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show

---


```{r}
library(dplyr)
library(readr)
library(ggplot2)

indicators <- read_csv("../Indicators.csv")
indicators <- select(indicators, CountryCode, IndicatorCode, Value, Year)
Countries <- read_csv("../Country.csv")
Countries <- select(Countries, CountryCode, ShortName)
Countries <- mutate(Countries, top_20 = ifelse(ShortName %in% c("Qatar", "Luxembourg", 
                                     "Singapore", "Brunei", "Kuwait", "Norway", "United Arab Emirates", "Hong Kong", "United States", "Switzerland", "Saudi Arabia", "Bahrain", "Netherlands", "Ireland", "Australia", "Austria", "Sweden", "Germany", "Taiwan", "Canada"), "Top 20 Richest Country", "Not Top 20"))
Countries <- mutate(Countries, world_power = ifelse(ShortName %in% c("China", "Japan", "Russia", "United States", "Germany", "France", "United Kingdom"), "World Power", "Not World Power"))
Countries
```

```{r}
#GDP (current LCU)
Variables <- merge(Countries, filter(indicators, IndicatorCode=="NY.GDP.MKTP.CD" &  Year==2012 & Value > 1e+3), by = 'CountryCode')
names(Variables)[names(Variables)=="Value"] <- "GDP"

#Birth rate (per 1,000 people)
Variables <- merge(Variables, filter(indicators, IndicatorCode=="SP.DYN.CBRT.IN" &  Year==2012), by = 'CountryCode')
names(Variables)[names(Variables)=="Value"] <- "Mortality"

GDPMortality <- Variables

#Central government debt, total (% of GDP)
Variables2 <- merge(Countries, filter(indicators, IndicatorCode=="GC.DOD.TOTL.GD.ZS" &  Year==2012), by = 'CountryCode')
names(Variables2)[names(Variables2)=="Value"] <- "Debt"

#Expense (% of GDP)
Variables2 <- merge(Variables2, filter(indicators, IndicatorCode=="GC.XPN.TOTL.GD.ZS" &  Year==2012), by = 'CountryCode')
names(Variables2)[names(Variables2)=="Value"] <- "Expense"


#countsCountry <- indicators %>%
#  group_by(CountryCode, CountryName) %>%
#  summarise(NumIndicators = n_distinct(IndicatorName),
#            NumYears      = n_distinct(Year),
##            FirstYear     = min(Year),
#            LastYear      = max(Year))

#countsIndicate <- indicators %>%
#  group_by(IndicatorCode, IndicatorName) %>%
#  summarise(NumCountries = n_distinct(CountryName),
#            NumYears     = n_distinct(Year),
#            FirstYear    = min(Year),
#            LastYear     = max(Year))

# Note - $ being replaced with dollar to avoid MathJax issue
#counts$IndicatorName <- gsub("\\$", "dollar", counts$IndicatorName)
#print.table(counts)
```

Graphings
```{r}
gdp_mortality_debt_allCountry <- indicators %>% 
  group_by(CountryName, Year, IndicatorName) %>%
  filter(IndicatorName == "GDP (current LCU)" |
         IndicatorName == "Mortality rate, under-5 (per 1,000)" | 
         IndicatorName == "Central government debt, total (% of GDP)" | 
         IndicatorName == "Military expenditure (% of GDP)")
```
```{r}

south_america <- gdp_mortality_debt_allCountry %>%
  filter(CountryName == "Brazil" |
         CountryName == "Argentina" | 
         CountryName == "Paraguay" | 
         CountryName == "Uruguay" |
         CountryName == "Chile" | 
         CountryName == "Bolivia" | 
         CountryName == "Peru" | 
         CountryName == "Ecuador" | 
         CountryName == "Colombia" | 
         CountryName == "Venezuela" | 
         CountryName == "Guyana" | 
         CountryName == "Suriname" )


gdpL <- filter(gdp_mortality_debt_allCountry, IndicatorName == "GDP (current LCU)")
nationalDebt <- filter(gdp_mortality_debt_allCountry, IndicatorName == "Central government debt (% of GDP)")

#gdpL_nationalDebt <- merge(gdpL, nationalDebt, by.x = "CountryName")
jnaegeli_315_theme <-  theme_bw() + 
  theme(
    axis.text = element_text(size = 10, family = "Palatino", 
                                 color = "royalblue4"),
        text = element_text(size = 14, family = "Palatino",
                            face = "bold", color = "grey30"))

#actually GDP^1/4 vs Birth rate (per 1,000 people)
ggplot(GDPMortality, aes(x = GDP ^ (1/4), y = Mortality, label = ShortName)) + geom_point(aes(color = world_power)) + 
  labs(
    title = "GDP vs Birth Rate",
    x = "(GDP)^(1/4) (current US Dollar)",
    y = "Birth Rate (per 1,000 people)",
    color = "World Power Status"
  ) + geom_smooth() + jnaegeli_315_theme + theme(legend.position="bottom")

#+
  #geom_text(data=subset(GDPMortality, GDP ^ (1/4) > 2000), check_overlap = TRUE)


ggplot(Variables2, aes(x = Expense, y = Debt, label = ShortName)) + geom_point(aes(color = world_power)) + 
  labs(
    title = "Government Expenses vs Government Debt",
    x = "Expenses (% of GDP)",
    y = "Central government debt (% of GDP)", 
    color = "World Power Status"
  ) + geom_smooth() +
  geom_text(data=subset(Variables2, Debt > 150)) + jnaegeli_315_theme + theme(legend.position="bottom")
#REMEMBER TO LABEL OUTLIER POINTS AND COLOR THE POINTS AS WELL

#ggplot(nationalDebt, aes(x = Year, y = Value)) + geom_point() + 
#  labs(
#    title = "Government Debt over the years",
#    x = "Central government debt total (% of GDP)",
#    y = "Year"
#  )

#ggplot(gdpL, aes(x = Year, y = Value)) + geom_point() + 
#  labs(
#    title = "GDP over the years",
#    x = "GDP (current LCU)",
#    y = "Year"
#  )
```


1st Graph:
The graph shows a country's GDP^(1/4) versus the birth rate. It seems that there is a general downward trend, as a Country's GDP rises, the birth rate will decrease. While the trend line may seem relatively horizontal, all of the world powers are under the trend line. World Powers are stronger, richer nations, so there is significance to the fact that all of them have low birth rates and are under the trend line. 

2nd Graph:
This graph shows a country's expenses versus its debt. As a country spends more, it owes more, increasing its debt. However, the more interesting details lie in the differences, and even lack of differences between the world powers and the rest of the countries. Most of the outliers and more extreme values come from non-world power nations. The only extreme value from a world power is Japan. Most of the world powers, while scattered from 0-100 debt-wise, tend to have lower expenses. However, it seems that without the extreme values, the top 20 nations are interspersed throughout the other nations without any visible group or correlation. This suggests that debt and expenses may play a smaller role in a country's wealth than previously thought. 


